Secure Composition of Robust and Optimising Compilers

### 2023-09-19 csf/main.tex
finally compiles

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[utf8,dvipsnames,conference,peerreviewca]{IEEEtran}

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{glossaries}
\usepackage{orcidlink}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}

@@

% https://tex.stackexchange.com/questions/648845/sans-serif-uppercase-greek-no-longer-showing-in-acmart
\DeclareMathAlphabet{\mathsf}{OT1}{LibertinusSans-LF}{m}{n}
\SetMathAlphabet{\mathsf}{bold}{OT1}{LibertinusSans-LF}{bx}{n}

\DeclareMathAlphabet{\mathtt}{OT1}{lmtt}{m}{n}
\SetMathAlphabet{\mathtt}{bold}{OT1}{lmtt}{bx}{n}

@@

@@

\begin{center}\small\it
	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
	% ~\citep{patrignani2020use}.
% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
	}
\end{center}
\end{abstract}

@@

\begin{example}[strncpy]\label{ex:strncpy}
Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
% Even under these assumptions, there is a 
This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
the memory location past the last element will be leaked to an attacker.
\begin{lstlisting}[language=c,basicstyle=\scriptsize\ttfamily,morekeywords={size_t}]
void strncpy(size_t n, char *dst, char *src) {
  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
    dst[i] = src[i];
  }
}
\end{lstlisting}

@@

Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
\end{example}

@@

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of~\citep{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
        % 
        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This is crucial for reordering optimisation passes and thus generating secure and efficient code.

@@

\begin{example}[Double Free in Bluetooth Subsystem]
  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
\end{example}

@@

\begin{example}[Types]
  Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
\end{example}

@@

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\hspace{0.5cm}
  \mi{(Event)}~\event\bnfdef\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}&~\event[_{b}]\bnfdef\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
  $$
  \mssafe:=\tmssafe \cap \smssafe
  $$
\end{definition}

@@

\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
  $$
  \msscctsafe:=\mssafe\cap\scctsafe
  $$
\end{definition}

@@

Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
%\MPpost{
%	this is fine, but it'd be nice to provide more insights on why it's easier
%}
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
\begin{figure}[!ht]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
\begin{center}
  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
  %
  \typerule{tms-refl}{
    \tmsmonitor.\text{allocated}=\emptyset
  }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
  }{tms-refl}
  %
  \typerule{tms-ign-trans}{
    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
  }{tms-ign-trans}
  %
  \typerule{tms-trans}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
  }{tms-trans}
\end{center}
\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
This rejects the behavior of programs that forget to free memory.

>>>>

% Fixing: Too many math alphabets used in version normal.
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[dvipsnames,conference]{IEEEtran}

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage[switch]{lineno}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{glossaries}
\usepackage{orcidlink}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}

@@

@@

\theoremstyle{definition}
\newtheorem{exampleenv}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]

@@

\begin{center}\small\it
	{This paper uses syntax highlighting accessible to both colourblind and black \& white readers.
	% ~\cite{patrignani2020use}.
% 	% Specifically, it makes use of a $\src{blue}$, $\src{sans\text{-}serif}$ font for a $\src{source}$,
% 	% a $\trg{red}$, $\trg{bold}$ font for an $\trg{intermediate}$,
% 	% and a $\obj{green}$, $\obj{teletype}$ font for a $\obj{target}$ language.
	}
\end{center}
\end{abstract}

@@

\begin{exampleenv}[strncpy]\label{ex:strncpy}
Consider the \texttt{C} function \texttt{strncpy} that copies a null-terminated string \texttt{src} into \texttt{dst} up to a length of \texttt{n}.
% Assume that the length of the array pointed by \texttt{dst} is at least \texttt{n}, and \texttt{src} is terminated by \texttt{'\textbackslash 0'}.
% Even under these assumptions, there is a 
This function is subject to a subtle \gls*{sms} vulnerability: the bounds check \texttt{i < n} should happen {\it before} the access to memory location \texttt{x[i]}: otherwise
% If \texttt{n} is at least as large as the length of the array \texttt{src}, 
the memory location past the last element will be leaked to an attacker.
\begin{lstlisting}[language=c,basicstyle=\scriptsize\ttfamily,morekeywords={size_t}]
void strncpy(size_t n, char *dst, char *src) {
  for(size_t i = 0; src[i] != '\0' && i < n; ++i) {
    dst[i] = src[i];
  }
}
\end{lstlisting}

@@

Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
\end{exampleenv}

@@

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
        % 
        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This is crucial for reordering optimisation passes and thus generating secure and efficient code.

@@

\begin{exampleenv}[Double Free in Bluetooth Subsystem]
  Consider \texttt{CVE-2021-3564}~\cite{doublefree-bluetooth}, one of many submissions for a double-free vulnerability.
  The vulnerability arises due to a race condition where the context-level function \texttt{hci\_cmd\_work} was not expected to behave maliciously, since it resides in the same source-code repository where the vulnerability occurs.
  Nevertheless, the component-level code of \texttt{hci\_dev\_do\_open} is linked with \texttt{hci\_cmd\_work} and does not atomically check whether a pointer has been freed already:
  Therefore, \texttt{hci\_dev\_do\_open} does not satisfy the no-double-frees property robustly, since there is an implementation for \texttt{hci\_cmd\_work} that leads to a violation of that property when linked with \texttt{hci\_dev\_do\_open}.
\end{exampleenv}

@@

% \begin{exampleenv}[Types]
%   Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
%   Consider the following $\src{L}$ component and its compiled version below.
%   \begin{center}
%   $
%   \begin{array}{l}
%     \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
%     \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
%   \end{array}
%   $
%   \end{center}
%   While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
%   Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
%   Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
%   The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
%   Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
% \end{exampleenv}

@@

\begin{gather*}
  \begin{aligned}
  \mi{(Security\ Tag)}~\sigma\bnfdef&\ \lock \mid \unlock\hspace{0.5cm}
  \mi{(Control\ Tag)}~\sandboxtag\bnfdef \ctx\mid\comp\\
  \mi{(Event)}~\event\bnfdef&\ \emptyevent \mid \lightning \mid \event[_{b}];\sandboxtag;\sigma \\
  \mi{(Pre\text{-}event)}~\event[_{b}]\bnfdef&\ \ev{Alloc\ \loc\ n} \mid \ev{Dealloc\ \loc} \\
                                             & \mid \ev{Use\ \loc\ n} \mid \ev{Branch\ n} \mid \ev{Binop\ n} \\
  \end{aligned}
\end{gather*}

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  \scriptsize$$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

\begin{definition}[\glsfirst*{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\tmssafe \cap \smssafe
  $
\end{definition}

@@

\begin{definition}[\glsfirst*{ms} and \gls*{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\mssafe\cap\scctsafe
  $
\end{definition}

@@

Monitors enforce safety properties by accepting or rejecting traces, i.e., if they reject a trace, the trace does not satisfy the property they check.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
%\MPpost{
%	this is fine, but it'd be nice to provide more insights on why it's easier
%}
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
\begin{figure}[!ht]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level\\trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsubsection{Monitor for \glsfirst*{tms}}\label{subsubsec:tms}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{A}:L\times\sandboxtag,\text{F}:L\times\sandboxtag\right\} \\%
    \emptyset:=&\ \left\{\text{A}:\emptyset,\text{F}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \\
                                            & \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$\\given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{A} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\cup\left\{(\loc;\sandboxtag)\right\},\text{F}: \tmsmonitor.\text{F}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{A} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{F} \\
    \tmsmonitor[']=\left\{\text{A}: \tmsmonitor.\text{A}\setminus\left\{(\loc;\sandboxtag)\right\}\right.,\\\left.\text{F}: \tmsmonitor.\text{F}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

For \gls*{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
Contrary to other monitors in this paper, the multi-step relation of the \gls*{tms} monitor is non-standard:
\begin{center}
  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$\\given trace $\abstrace$.''}$\;$\\
  %
  \typerule{tms-refl}{
    \tmsmonitor.\text{allocated}=\emptyset
  }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
  }{tms-refl}
  %
  \typerule{tms-ign-trans}{
    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
  }{tms-ign-trans}
  %
  \typerule{tms-trans}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
  }{tms-trans}
\end{center}
\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
This rejects the behavior of programs that forget to free memory.

\begin{center}
  \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$\\with respect to \gls*{tms}.''}$\;$\\
  %
  \typerule{tms-alloc-authentic}{
  }{
    \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc-authentic}
  %
% \typerule{tms-dealloc-authentic}{
% }{
%   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
% }{tms-dealloc-authentic}
  %
% \typerule{tms-use-authentic}{
% }{
%   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
% }{tms-use-authentic}
  %
  \typerule{tms-branch-authentic}{
  }{
    \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
  }{tms-branch-authentic}
  %
% \typerule{tms-binop-authentic}{
% }{
%   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
% }{tms-binop-authentic}
% %
% \typerule{tms-empty-authentic}{
% }{
%   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
% }{tms-empty-authentic}
% %
  \typerule{tms-abort-authentic}{
  }{
    \tmstraceagree[]{\lightning}{\absterm}
  }{tms-abort-authentic}
\end{center}
